Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-17 paper.qmd
Added function to plot the 2D projection

@@

pbmc_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    #size = 0.5,
    alpha = 0.2,
    color = '#a65628') +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("a2", cex = 2)

@@

pbmc_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2)) +
  geom_segment(
    data = model_df,
    aes(
      x = proj1_from,
      y = proj2_from,
      xend = proj1_to,
      yend = proj2_to),
    color = "#000000") +
  geom_point(
    #size = 0.5,
    alpha = 0.2,
    color = '#a65628') +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2),
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 0.75)) +
  ylim(c(-1, 0.75)) +
  interior_annotation("a3", cex = 2)

@@

@@

mnist_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5,
    alpha = 0.5) +
  geom_point(
    size = 0.5,
    alpha = 0.5) +
  scale_color_manual(values = c("#999999",'#ff7f00')) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  #coord_fixed() +
  xlim(c(-1.1, 0.9)) +
  ylim(c(-1.1, 0.9)) +
  interior_annotation("b", cex = 1.5) +
  theme(
    legend.position = "none"
  )

>>>>

plot_proj <- function(projected_df, model_df, axes, circle, 
                      point_param = c(1.5, 0.5, "#000000"), # size, alpha, color
                      line_param = c(0.5, 0.5), #linewidth, alpha
                      plot_limits, title, cex = 2, 
                      axis_text_size = 3,
                      is_category = FALSE) {

  if(is_category == FALSE) {

    initial_plot <- projected_df |>
      ggplot(
        aes(
          x = proj1, 
          y = proj2))

  } else {

    initial_plot <- projected_df |>
      ggplot(
        aes(
          x = proj1, 
          y = proj2,
          colour = cluster))

  initial_plot + 
    geom_segment(
      data = model_df, 
      aes(
        x = proj1_from, 
        y = proj2_from, 
        xend = proj1_to, 
        yend = proj2_to), 
      color = "#000000",
      linewidth = line_param[1],
      alpha = line_param[2]) +
    geom_point(
      size = as.numeric(point_param[1]),
      alpha = as.numeric(point_param[2]),
      color = point_param[3]) +
    geom_segment(
      data=axes, 
      aes(x=x1, y=y1, xend=x2, yend=y2), 
      colour="grey70") +
    geom_text(
      data=axes,
      aes(x=x2, y=y2), 
      label=rownames(axes),
      colour="grey50",
      size = axis_text_size) +
    geom_path(
      data=circle, 
      aes(x=c1, y=c2), colour="grey70") +
    xlim(plot_limits) +
    ylim(plot_limits) +
    interior_annotation(title, cex = cex)

@@

pbmc_proj_umap_model1 <- plot_proj(
  projected_df = projected_df, 
  model_df = model_df, 
  axes = axes, 
  circle = circle, 
  point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
  plot_limits = c(-0.65, 0.65), 
  title = "a2", 
  cex = 2, 
  is_category = FALSE)

@@

pbmc_proj_umap_model2 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
    plot_limits = c(-1, 0.75), 
    title = "a3", 
    cex = 2, 
    is_category = FALSE)

@@

tr_from_to_df_mnist <- inner_join(
  tr_from_to_df_mnist, distance_df_small_edges_mnist, 
  by = c("from", "to"))

@@

mnist_proj_tsne_model1 <- plot_proj(
  projected_df = projected_df, 
  model_df = model_df, 
  axes = axes, 
  circle = circle, 
  point_param = c(0.5, 0.5, "#000000"), # size, alpha, color
  line_param = c(0.5, 0.5), # linewidth, alpha
  plot_limits = c(-1.1, 0.9), 
  title = "b", 
  cex = 2, 
  axis_text_size = 3, 
  is_category = TRUE) +
  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
  theme(
    legend.position = "none"
  )
